The Effectiveness of Strategies to Contain SARS-CoV-2: Testing, Vaccinations, and NPIs

### 2022-03-10 src/rev1_reply_comments.tex
Answers to ref 2.

\begin{enumerate}
    \item They use contact-frequency from pre-pandemic diary data. Is  this truly useful
          if behavior of people is already changed by the fact there is a pandemic?
          \response{%

@@

              Furthermore, note that we only use the pre-pandemic diary data as the
              starting point. Contact restrictions and individual behavior adjustments
              reduce contacts. These are active throughout the period of analysis and
              they vary with the strictness of measure. In order to do so, it is crucial
              to take the type of contact into account; see for example our analysis of
              potential school closures on Page~\pageref{ref.2.a}. Closed schools mean
              that children have far fewer and parents have somewhat fewer contacts than
              at  and for the most part not randomly but in very specific ways. This
              allows us to account for behavior reactions and policy effects that target
              different types of contacts differently. For example, household contacts
              continue to take place no matter the policy environment. Work from home
              does not keep random individuals at home but every individual that works
              has a time-constant value that determines when he or she will stay home.
              Which share of the population stays home is taken from Google mobility
              data. In the education sector we do not use the Mossong data but use
              completely mechanistic models. This gives us confidence that we can
              sufficiently accurately represent the contact patterns during the pandemic
              and account for changes induced by the pandemic.

              Also, our excellent fit for most age groups gives us confidence that we
              manage to match actual meeting patterns. If we were far off for example
              for the number of work contacts, we would not be able to match either the
              infections of individuals in the working age (35-59) or of those that
              mostly do not work anymore (60-79).

              \textbf{Answer in case there is confusion because the other multiplier is 1 in fall 2020:}

              We account for individual responses in behaviour to the pandemic through a
              time-variant multiplier that reduces the number of other (i.e. non-work,
              non-household, non-educational) contacts. The values at change points that
              mark policy announcements or fatigue effects are estimated. They are
              detailed in Section A.9.

              Since we estimate both the infection probability of this contact type and
              the multiplier, one of them has to be normalised. We decided to normalise the
              multiplier to 1 in October 2020. This may appear as if we assumed that
              pre-pandemic behaviour had resumed during that time in Germany. However, this
              is not the case. In fact, we would achieve the exact same fit if we doubled
              the infection probability and halved the multiplier. This is because halving
              the multiplier means only half the number of contacts take place. However,
              because of the doubling of the infection probability, in expectation we would
              arrive at the same number of infections in that contact category. }
    \item They state that susceptibility is dependent on age. This requires references
          and needs to be further explained. Also, younger populations are highly susceptible,
          they may just not contribute transmission as much. On that same note, infectiousness
          may not be independent of age

@@

          \response{%
              We now cite our main reference for the susceptibilities for each age group
              also in the main text. We directly take the susceptibilities given by
              Davies et al. 2020, in their Extended Data Fig. 4. We base our assumption
              on infectiousness being independent of age on Jones et al. 2021. They
              report only slight differences in viral loads between German adults and
              children: ``The least infectious youngest children have 78\% (61, 94) of
              the peak culture probability of adults aged 45 to 55''. Furthermore, they
              caution that these lower numbers may be due to different sample taking -
              such as the usage of smaller pediatric swab devices - rather than actual
              lower viral loads and conclude that the differences, if existent, are
              likely not clinically relevant. }
    \item It is not accurate that the vaccines stop transmission. This has to be adapted
          accordingly. Also, the authors should state which vaccines they based this
          analysis on, as in which ones were in use in Germany at that time and how does
          that affect the relative transmissibility as it cannot be assumed that all
          vaccines will have the same effect. \response{%
              We agree that vaccines do not stop transmission completely. We simplify
              the way vaccines reduce transmission risk considerably by assuming that
              75\% of individuals achieve sterile immunity through a vaccination while
              the remaining 25\% remain fully susceptible. We are comfortable with this
              assumption for several reasons: Firstly, the existing evidence supports
              the assumption that all four vaccines approved in the EU (Biontech/Pfizer,
              Moderna, AstraZeneca and Johnson \& Johnson) provide similar levels of
              protection in the first few months after vaccination and against the Alpha
              variant (for details see the last third of appendix A.1), which is the
              only time period and variant our simulations cover. Secondly, given the
              surprisingly good performance of rapid tests we tried to err on the side
              of being too optimistic with our modelling of vaccine performance. This is
              why we chose the perfect protection from vaccination for a large part of
              the population. Thirdly, reinfections are not a big issue for the time
              period that we look at. The incidences were small and the chance of more
              than one exposure to Covid are negligibly small. Thus, the value added by
              a probabilistic model of vaccination protection is very small for the
              period that we cover. }
\end{enumerate}
}

>>>>

\begin{enumerate}
    \item They use contact-frequency from pre-pandemic diary data. Is  this truly useful
          if behavior of people is already changed by the fact there is a pandemic?

          \response{%

@@

              Furthermore, note that we only use the pre-pandemic diary data as the
              starting point. Contact restrictions and individual behavior adjustments
              reduce contacts. These are active throughout the period of analysis and
              they vary with the strictness of measure. In order to do so, it is crucial
              to take the type of contact into account; see for example our analysis of
              potential school closures on Page~\pageref{ref.2.a}. Closed schools mean
              that children have far fewer and parents have somewhat fewer contacts than
              without school closures.

              Finally, the very good fit of our model to official infection rates across
              age groups increases our confidence that we manage to match actual meeting
              patterns. If we were far off for, for example, the number of work contacts,
              we would not be able to match either the infections of individuals in the
              working age (35-59) or of those that mostly do not work anymore (60-79).
          }
    \item They state that susceptibility is dependent on age. This requires references
          and needs to be further explained. Also, younger populations are highly susceptible,
          they may just not contribute transmission as much. On that same note, infectiousness
          may not be independent of age

@@

          \response{%
              Many thanks for pointing this out, indeed we had not cited our key reference for susceptibilities by age group \citep{Davies2020} in the main text. We do so now on Page~\ref{r.2.b}.

              We base our assumption on infectiousness being independent of age on
              \citep{Jones2021}. That paper reports only slight differences in viral
              loads between German adults and children: ``The least infectious youngest
              children have 78\% (61, 94) of the peak culture probability of adults aged
              45 to 55''. Furthermore, they caution that these lower numbers may be due
              to different sample taking---such as the usage of smaller pediatric swab
              devices---rather than actual lower viral loads and conclude that the
              differences, if existent, are likely not clinically relevant.

                  {\bf ToDo: Check paper suggestion
                      \url{https://doi.org/10.1093/infdis/jiaa691} -- are they actually checking
                      infectiousness conditional on contact type? Or could we even reproduce
                      what they are doing (not suggesting we should, certainly no need to spend
                      time on it)}
              %   In any case, while the jury is certainly still out on that particular issue, we are confident that different assumptions would not change our results much. 
          }
    \item It is not accurate that the vaccines stop transmission. This has to be adapted
          accordingly. Also, the authors should state which vaccines they based this
          analysis on, as in which ones were in use in Germany at that time and how does
          that affect the relative transmissibility as it cannot be assumed that all
          vaccines will have the same effect.

          \response{%
              Many thanks, it is of course true that vaccines do not stop transmission
              completely. We do not assume that, either. However, we have not been able
              to find data on the way in which this affects different population groups
              for the wild type and the $\alpha$-variant.

              In our model, we thus simplify the way vaccines reduce transmission risk.
              In particular, we assume that that 75\% of individuals achieve sterile
              immunity through a vaccination while the remaining 25\% remain fully
              susceptible. SFirstly, the existing evidence supports the
              assumption that all four vaccines approved in the EU (Biontech/Pfizer,
              Moderna, AstraZeneca and Johnson \& Johnson) provide similar levels of
              protection in the first few months after vaccination against the wild type
              and the $\alpha$ variant (for details see Appendix~\ref{r.2.c}, starting
              on Page~\pageref{r.2.c}), which is the only time period and variant our
              simulations cover. Secondly, given the surprisingly good performance of
              rapid tests we tried to err on the side of being too optimistic with our
              modelling of vaccine performance. This is why we chose the perfect
              protection from vaccination for a large part of the population. Thirdly,
              reinfections are not a big issue for the time period that we look at. The
              incidences were small and the chance of more than one exposure to Covid
              are negligibly small. Thus, the value added by a probabilistic model of
              vaccination protection is very small for the period that we cover. }
\end{enumerate}
}
